Towards Artificial Intelligence for Metaphysics: \\ the Case of the Inconsistency in G\"odel's Ontological Argument

### 2016-01-07 Papers/InProgress/IJCAI/paper.tex
Add packages and commands
Minor changes

% the following package is optional:
%\usepackage{latexsym}

\usepackage{graphicx,url}

@@

\begin{itemize}
\item A detailed analysis of the inconsistency of G\"{o}del's
  \shortcite{GoedelNotes} original version of the proof is presented
  for higher-order modal logics KB and K (with possibilist and
  actualist first-order quantifiers). The extraction, reconstruction
  and verification of an informal, human intuitive argument has been
  an open problem since the first detection of this inconsistency by
  Benzmüller and Woltzenlogel-Paleo \shortcite{C40} with the LEO-II prover. 
  The experiments reported here led to the discovery of a surprisingly accessible
  understanding of this inconsistency which is philosophically
  profound and which has not been presented yet in the
  literature. The detection of this inconsistency in combination with
  the work reported here thus demonstrates that artificial intelligence systems -- 
  particularly higher-order automated
  theorem provers -- are capable of assisting in the discovery and ellucidation of
  \emph{new} and philosophically relevant knowledge. 
\item Another open question has been whether the final theorem T3 \textit{``Necessarily, there
    exists God''} could eventually be proved fully automatically
  directly from the axioms alone. In \shortcite{C40}, the provers had only succeeded in proving 
  T3 from intermediate argumentation steps (lemmata). Here we report that, by using a more efficient new
  embedding for the higher-order modal logic S5, T3 can be directly proven from Scott's version of the axioms, which are consistent. The fully automatic proof has been
  generated by the theorem prover \textsc{Leo-II}~\cite{C26} and
  subsequently verified in the proof assistant
  Isabelle/HOL~\cite{NPW02}.
\item A third open challenge the verification in Isabelle/HOL of the modal
  collapse \cite{Sobel}, which is one of the most strongly criticised
  `side-effects' of G\"odel's and Scott's variants of the proof. In previous work
  the modal collapse has been derived by the provers
  \textsc{Satallax} \cite{Satallax} and \textsc{Leo-II}, but a fully automatic
  verification in the highly trusted Isabelle/HOL still failed. 
  Now, with the more efficient embedding for S5, this verification has been done. 
\item On the technical side, we present a significantly improved
  syntax representation of G\"odel's resp.~Scott's proof in
  Isabelle/HOL. In fact, we demonstrate that a nearly perfect match
  between the original pen and paper presentations and our encoding in
  Isabelle/HOl is feasible. This is clearly an important prerequisite
  for promoting the theorem proving technology employed here to a wide community of
  philosophers, who are not necessarily experts in automated reasoning or higher-order logic.
\item On the other hand, we
  also point to several technical challenges which, at least for the
  moment, still restrain non-expert users of these systems from
  independently conducting similar experiments in computational
  metaphysics as reported here and in previous work. 
% Technical issues detected within Isabelle-Sledgehammer; room for improvements
\end{itemize}

% These novel contributions are presented in Sections ? and ?.
% In Section ?, novel and previous findings are summarised and discussed 
% with regard to their relevance in metaphysics. In this section also 
% some further details and improvements on earlier 
% findings are presented.
% Section ? concludes the paper.

Give paper structure (maybe not needed) ...

>>>>

% the following package is optional:
\usepackage{latexsym}

\usepackage{amsmath}
\usepackage{modallogics}
\usepackage{graphicx,url}

\newcommand{\imp}{\rightarrow}
\newcommand{\biimp}{\leftrightarrow}
\newcommand{\allq}{\forall}
\newcommand{\exq}{\exists}
\newcommand{\seq}{\vdash}
\newcommand{\nec}{\Box} % necessarily
\newcommand{\pos}{\Diamond} % possibly

@@

\begin{itemize}
\item A detailed analysis of the inconsistency of G\"{o}del's
  \shortcite{GoedelNotes} original version of the proof is presented
  for higher-order modal logics KB and K (with possibilist and
  actualist first-order quantifiers). The extraction, reconstruction
  and verification of an informal, human intuitive argument has been
  an open problem since the first detection of this inconsistency by
  Benzm\"uller and Woltzenlogel-Paleo \shortcite{C40} with the LEO-II prover. 
  The experiments reported here led to the discovery of a surprisingly accessible
  understanding of this inconsistency which is philosophically
  profound and which has not been presented yet in the
  literature. The detection of this inconsistency in combination with
  the work reported here thus demonstrates that artificial intelligence systems -- 
  particularly higher-order automated
  theorem provers -- are capable of assisting in the discovery and ellucidation of
  \emph{new} and philosophically relevant knowledge. 
\item Another open question has been whether the final theorem T3 \textit{``Necessarily, there
    exists God''} could eventually be proved fully automatically
  directly from the axioms alone. In \shortcite{C40}, the provers had only succeeded in proving 
  T3 from intermediate argumentation steps (lemmata). Here we report that, by using a more efficient new
  embedding for the higher-order modal logic S5, T3 can be directly proven from Scott's version of the axioms, which are consistent. The fully automatic proof has been
  generated by the theorem prover \textsc{Leo-II}~\cite{C26} and
  subsequently verified in the proof assistant
  Isabelle/HOL~\cite{NPW02}.
\item A third open challenge the verification in Isabelle/HOL of the modal
  collapse \cite{Sobel}, which is one of the most strongly criticised
  `side-effects' of G\"odel's and Scott's variants of the proof. In previous work
  the modal collapse has been derived by the provers
  \textsc{Satallax} \cite{Satallax} and \textsc{Leo-II}, but a fully automatic
  verification in the highly trusted Isabelle/HOL still failed. 
  Now, with the more efficient embedding for S5, this verification has been done. 
\item On the technical side, we present a significantly improved
  syntax representation of G\"odel's resp.~Scott's proof in
  Isabelle/HOL. In fact, we demonstrate that a nearly perfect match
  between the original pen and paper presentations and our encoding in
  Isabelle/HOl is feasible. This is clearly an important prerequisite
  for promoting the theorem proving technology employed here to a wide community of
  philosophers, who are not necessarily experts in automated reasoning or higher-order logic.
\item On the other hand, we
  also point to several technical challenges which, at least for the
  moment, still restrain non-expert users of these systems from
  independently conducting similar experiments in computational
  metaphysics as reported here and in previous work. 
% Technical issues detected within Isabelle-Sledgehammer; room for improvements
\end{itemize}
